---
title: "comm_metricsNWT"
author: ""
date: "01 March 2019"
output: pdf_document
---

# Overview

The **comm_metricsNWT** module computes a series of community metrics based on the bird predictions for the Northwestern Territories (NWT) ('birdsNWT' module).

The computed metrics are:
 * Expected Species Richness*: it is considered as probability of occupancy based on a Poisson distribution. 
 $$ S: 1 - \exp^{\lambda}$$  
 where:  $$ exp^{-\lambda} = \frac{exp^{-\lambda}\lambda^k{}{k!} $$
 
 * Shannon diversity Index*: 
 $$ H^' = - \sum_{i=1}^{S} p_ilnp_i $$

where:$$p_i$$ is the proportion of individuals belonging to the ith species.

 * Simpson diversity Index*: 
 $$ 1-D = \sum p_i^2 $$

 * Rao's Quadratic Entropy: $$ \sum_{i=1}^{S-1} \sum_{j=i+1}^{S} d_ij* p_i* p_i$$

 Raos Quadratic Entropy incorporates two elements: 1)relative abundances of species and 2)a measure of pairwise functional differences between species [^1]. For this second element, we used *Life history traits* as considered in Knaggs (2018)[^2]. 
 The life history traits were originally produced by the Avian Life History Information Database Wildspace support system (http://www.on.ec.gc.ca/wildlife/wildspace/project.cfm ) but they are not longer available online. 


**References**
[^1] Botta-Dukát, Z.2005. Rao's quadratic entropy as a measure of functional diversity based on multiple traits. Journal of Vegetation Science 16:533-540. 

[^2] Knagss, M. 2018. Effects of burn severity and time since fire on songbird communities in the northern boreal forest. Department of Renewable Resources. University of Alberta. Master's Thesis. 79 pp.

# Usage

```{r module_usage}
library("SpaDES")
library("raster")
library ("rgdal")
library ("sp") 

# Source functions in R folder
invisible(sapply(X = list.files(file.path(getwd(), "R"), full.names = TRUE), FUN = source))

# Set a storage project folder
setPaths(modulePath = file.path("C:/Users/AARAS2/Downloads/NWT/modules"))
paths <- getPaths() # shows where the 4 relevant paths are


times <- list(start = 0, end = 3)

parameters <- list(
  .progress = list(type = "text", interval = 1), # for a progress bar
  ## If there are further modules, each can have its own set of parameters:
birdsNWT = list(
    "baseLayer" = 2005,
    "overwritePredictions" = TRUE,
    "useTestSpeciesLayers" = TRUE, # Set it to false when you actually have results from LandR_Biomass simulations to run it with
    "useParallel" = FALSE, # Using parallel in windows is currently not working.
    "predictionInterval" = 1
  )
)
.objects <- list(
  "birdsList" = c("BBWA", "BOCH","RBNU"))
modules <- list("birdsNWT", "comm_metricsNWT")
inputs <- list()
outputs <- list()
comm_metricsNWT <- simInitAndSpades(times = times, params = parameters, modules = modules,
                 objects = .objects, loadOrder = c("birdsNWT","comm_metricsNWT"), debug = 2)

```

The input data is the birdPrediction list produce by the module "birdsNWT".
A list of Life history traits for each bird species (sp_traitsNWT)

This module produces rasters for each diversity metric for the current year. The results can be access by:
```{r results}
comm_metricsNWT$currentDiversityRasters
```

In addition, this module produces a two tables with the mean value for each metric. The results can be access by:
```{r stats}
comm_metricsNWT$diversitystats
comm_metricsNWT$diversityByPolygon
```